\documentclass[10pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry}
\usepackage{graphicx}

% page layout issues
\setlength{\textwidth}{12.25cm}
\setlength{\oddsidemargin}{0.5\paperwidth}\addtolength{\oddsidemargin}{-2.54cm}\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\evensidemargin}{\oddsidemargin}
\setlength{\textheight}{22.86cm}
\setlength{\topmargin}{0.5\paperheight}\addtolength{\topmargin}{-\headheight}\addtolength{\topmargin}{-\headsep}\addtolength{\topmargin}{-2.54cm}\addtolength{\topmargin}{-0.5\textheight}
\renewcommand{\MakeUppercase}[1]{#1} % hacking the latex margin defaults; this is a bad hack

% text typesetting and font issues
\renewcommand{\paragraph}[1]{\bigskip\noindent\textbf{#1}~---}
\renewcommand{\subparagraph}[1]{\medskip\noindent\textsl{#1}~---}
\newcommand{\foreign}[1]{{\slshape #1}}
\sloppy\sloppypar\raggedbottom\frenchspacing

% text macros
\newcommand{\chapterref}[1]{\chaptername~\ref{#1}}

% math macros
\newcommand{\given}{\,|\,}

\title{\bfseries%
The Theory and Practice of Data Analysis
for the Physical Sciences}
\author{David W. Hogg}
\date{July 2025}

\begin{document}

\maketitle
\tableofcontents

\include{preface}

\chapter*{Subjects that need a home}\addcontentsline{toc}{chapter}{Subjects that need a home}\markboth{Subjects that need a home}{Subjects that need a home}
This placeholder chapter is a place to put subjects that ought to go somewhere in the book, but I don't know where.
As I figure out where, I will move these ideas to those places.
\begin{itemize}
  \item The significance of your data is not itself significant.
  Consequence: Don't cut on SNR or significance; cut on first-order properties.
  Related: You can't know second-order things about your data as well as you know first-order things.
  Cut on first-order things in such a way that you mimic the SNR cut you have in mind, instead.
  \item How about upper limits vs detections?
  Related to forced photometry and catalog matching.
  \item Catalog matching? Is this too astronomical?
  \item \textsl{[Add more things above here.]}
\end{itemize}

\chapter{How to structure a data analysis project}

\part{Basics}

\chapter{Fitting a model to data}

\include{likelihood}

\include{measurement}

\chapter{Estimating uncertainty}

\chapter{Model selection or decision making}

\chapter{Predicting and betting}

\chapter{Describing your model to others}\label{ch:notation}

\part{Theory}

\chapter{Probability theory for data analysis}

\chapter{Information theory}

\chapter{Should a project be frequentist or Bayesian?}
For insertion somewhere in this \chaptername.
\begin{itemize}
\item The idea that a prior can be because you think that the variable is actually, in the world, distributed like the prior.
Or it can be because you have ignorance about the variable, and that ignorance has a probabilistic form.
For example, angles can be uniformly distributed because a system is angle-mixed, or because \emph{you are not observing it at any special time}.
\item The idea that the prior on a nuisance parameter can be literally \emph{part of your model}. for example, when a vector is isotropically distributed (like an angular momentum vector for planetary systems), then you know the distribution of this parameter and you can integrate it out, even if you are a frequentist.
It becomes like a noise source.
\end{itemize}

\chapter{Choosing and testing prior pdfs}

\chapter{Causal structure and causal inference}

\chapter{Very small samples and \foreign{a posteriori} statistics}

\chapter{Exact symmetries}\label{ch:symmetry}

\chapter{Likelihood-free inference}

\part{Practice}

\chapter{Instrument calibration}

\chapter{Selection functions and censored data}

\chapter{Interpolation and very flexible models}\label{ch:flexible}

\chapter{Gaussian processes}

\chapter{Markov Chain Monte Carlo}

\chapter{Regression, including machine-learning regressions}

\chapter{Missing and heteroskedastic data}

\chapter{Robust methods and outliers}

\chapter{Model checking and visualization}

\chapter{Writing a scientific paper}

\appendix
\part*{Appendices}\addcontentsline{toc}{part}{Appendices}\markboth{}{}

\chapter{Useful probability distributions}

\chapter{Properties of Gaussian distributions}

\chapter{Data management}

\chapter{Numerical precision and code practices; testing}

\end{document}
